---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80
```{r}
set.seed(42)
population <- tibble (
  age = sample(18:80, 1000, replace = TRUE)
)

head(population)
# fortnite
```


### Task 3: Sample 100 rows of data from the dataset
```{r}
sample_100 <-population |> 
  sample_n(100)

sample_50 <- population |> 
  sample_n(50)

sample_10 <- population |> 
  sample_n(10)
```


### Task 4: Calculate the mean age and standard deviation of the sample
```{r}
sample_100 |> summarize(mean = mean(age), sd = sd(age))
sample_50 |> summarize(mean = mean(age), sd = sd(age))
sample_10 |> summarize(mean = mean(age), sd = sd(age))
```


### Task 5: Calculate the mean age and standard deviation of the population
```{r}
population |> summarize(mean = mean(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

Yeah they're very similar.Even going down from 100 ages to 50, the data was still pretty close. We had to shrink it down to 1% of the population to see real shifts in mean. Interesting to me how small the sample size could be while still being accurate.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
