---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse
grah
```{r}
library(tidyverse)
```



### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv
```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```



### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (`Fall 2024` - `Fall 2021`)/`Fall 2021` * 100) 
```


### Task 4: Describe our results

Write a few sentences about what you found in the data. What's the story here?
150% increase in fees for Student Sustainability. Silent increases in these fees in this area as well
as in Student Counseling are inherently toxic, in that the university decided to raise these fees by small
margins that would normally fly under the radar. 0% change in Athletics fees, which will change next year 
due to UMD finally paying their athletes.




When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
