# JOUR405: Statistics for Journalists
## Midterm Exam - Spring 2025

Name: Joseph Conaty

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences. Start by loading the tidyverse and any other libraries you think you might need.

```{r}
library(tidyverse)
```

## Part 1: Restaurant Health Inspections (15 points)

You want to understand how restaurants in Montgomery County are performing on health inspections. The first dataset contains restaurant health inspection scores for restaurants in Montgomery County. The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv` and complete these tasks:

### Tasks:
1. Calculate the mean and standard deviation of compliance scores (5 points)
2. Create a histogram of the compliance scores with a vertical line showing the mean (5 points)
3. Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? (5 points).

```{r}
restauraunts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```
Tasks!!!
```{r}
restauraunts |> summarize(mean = mean(compliance_score), sd = sd(compliance_score))
```
```{r}
restauraunts |> 
  ggplot()+
  geom_histogram(aes(x = compliance_score), binwidth = 5)+
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", linewidth = 1)
```
Ok so the mean compliance score is roughly a 96. With a deviation of slightly higher and slightly lower. This means that restauraunts in Mogo are mostly compliant with health inspection policies! So that is newsworthy for people in the county to know the places they eat are healthy! I think.

## Part 2: High School Athletics (25 points)

You are reporting a story about high school sports participation in Maryland and want to see if there are differences between boys and girls. The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and sex. Load the data from: `https://raw.githubusercontent.com/example/md_hs_sports_2024.csv` and complete these tasks:

### Tasks:
1. Calculate the correlation between boys' and girls' participation (5 points)
2. Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. (5 points)
3. Create a scatterplot showing this relationship, adding a line of best fit (5 points)
4. In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

```{r}
athletics <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```
```{r}
athletics |> summarize(correlation = cor (boys, girls))
```

```{r}
total <- athletics |> mutate(total = boys + girls)
girls_pct <- total |> mutate(girls_pct = (girls/total))
```

```{r}
total |> 
  ggplot()+
  geom_point(aes(x =boys, y = girls))+
  geom_smooth(aes(x =boys, y = girls), method = "lm", binwidth = 100)
```
Sentences:
Correlation revealed a very strong relationship between boy and girl participation at 0.98. The scatterplot also showed a pretty strong relationship but it wasn't what I expected it to be, kind of spread out. Maybe it's because of the x and y values being so high. I think districts below the line aren't encouraging girls to try out for sports or giving them as many opportunities to play in sports. Districts above the line are giving multiple opportunities for girls to play sports, but both are giving boys opportunities to play sports.
## Part 3: Public Transit Ridership (20 points)

You are investigating public transit ridership in the Washington, D.C. area and want to understand the patterns of daily bus and rail ridership. The third dataset contains daily bus and rail ridership totals from WMATA for the past year. Load the data from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv and do the following:

### Tasks:
1. Calculate the average bus and rail ridership and standard deviation using summarize() (5 points)
2. Using the process you used in class, take a random sample daily ridership numbers and calculate the sample means and deviations for bus and rail. The number in the sample is up to you, but explain why you chose what you did. Compare this to the stats you generated in step 1. (5 points)
3. Using group_by() and summarize(), calculate the means for bus and rail ridership for each weekday. Describe the overall pattern of ridership for bus and rail - which days stand out and why? Are there differences between bus and rail in the standard deviation values? (10 points)


```{r}
bus_rail <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```
```{r}
bus_rail |> summarize(sd = sd(bus), sd(rail))
bus_rail |> summarize(mean = mean(bus), mean(rail))
```
```{r}
bus_sample <- bus_rail |> 
  sample_n(100)
rail_sample <- bus_rail |> 
  sample_n(200)
```
```{r}
bus_sample |> summarize(sd = sd(bus), sd(rail))
rail_sample |> summarize(mean = mean(bus), mean(rail))
```
Chose 100 and 200 because I saw there were like a lot of days. And if I sampled 10 would that really be enough to cover hundreds of days accurately? I figured it would be more accurate to do 100 and 200 just to get a solid range.
```{r}
bus_rail |>  summarize(mean = mean(bus), mean(rail))
```



## Part 4: Maryland Car Theft Rates (20 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

### Tasks:
1. Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate (5 points)
2. Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median, and what percentage of all car thefts occur in those counties? (5 points)
3. Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? (10 points)


```{r}
car_theft_rates <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```
```{r}
county_rate <- car_theft_rates |> mutate(county_rate = `2023`/population*10000)
```

```{r}
median_rate <- county_rate |> summarize(median = median(county_rate))
```

```{r}
total_thefts <- car_theft_rates |> summarize(sum = sum(`2023`))
```
Calculations reveal that car thefts in MD happen a lot more in Baltimore City and PG County than in pretty much every other county. Rates are much higher in these two counties than the median rate of 9.5 for counties in general. And as for total, the total thefts in 2023 was like 23,000. And just in PG and Bmore it was roughly 16,000 car thefts. They made up for 16,000 / 23,000 car thefts in Maryland.I think that's the lede.
## Part 5: Data Analysis Scenario (20 points)

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences (no code!) explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)
  I would calculate mean emergency response times this year and last year. Then I would calcualte the SD of emergency response times for this and last year. But importantly, I would separate the calculations between police, fire, and ambulance calls since combining those categories is intellectually dishonest. 
2. What visualizations would help readers understand the trends? (5 points)
  I think a histogram with a mean line would help readers see if response times have gotten much lower like the tip says. If it's true, readers would see a big red line that shows what they were at and then all the new data on the left side of it.
3. What additional context or data would you need to make this a complete story? (5 points)
  I would need to ask people who work at these call centers to see if they've had to adjust to new technology, which may affect how fast they respond to calls. Another is to see how weather events as well as layoffs skew the data.


### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Before submitting, clear your environment and run the entire notebook

Good luck!
