---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

ANSWER HERE
The more important categories in the data set are homicides and sexual offences as well as robberies. Traffic accidents, although more frequent than most of the other categories happen and although they can lead to fatalities it is mostly just some car damage and people move on. It's hard to see any linear pattern in the data. And also I'm not really sure what the difference is between auto theft and theft from auto. That makes it a bit hard to articulate what I am taking away from the data. Maybe that auto theft and traffic accidents are high in PG, maybe that means we need to teach people how to drive better in the county.

### Task 4: Calculating Averages

Calculate the total number of incidents for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
total_incidents <- pg_crime |> mutate(total_incidents = (Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec))

pg_crime_averages <- pg_crime |> mutate(average_incidents = (Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec)/12)
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

ANSWER HERE
The categories with the highest incidents per month are traffic accidents, thefts, stolen cars, and thefts from cars. The lowest ones are are assaults with weapons, residential robberies, sexual offences, homicides, and non commercial breaking and entering. This tells me that an average month in PG county is filled with a lot of car related incidents and crimes. I guess a question I have is of the different between "theft from auto" and "auto theft" cause I don't know if I understand it or not. Is that stealing from somebody while you're in a car and they're not in a car? Auto theft would be stealing someone's car so I assume that's what it means.

### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(june_change = (Jun-average_incidents), july_change = (Jul-average_incidents), august_change = Aug - (average_incidents))
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

ANSWER HERE
Perchange change in avg incidents per month was only really completely positive for car related incidents. Car theft, other vehicle theft, car thefts that were recovered, and theft from cars. So that's to say that the perception of crime being higher in the summer is only true in terms of car related crimes. Most other crimes had negative percent changes during the summer, except for homicides which were only slightly positive because they were so low in the first place compared to other data. 

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE
PG County car theft is highest during the summer, according to annual PG Crime data.

Being in a car at all makes you extremely likely to be part of this incident stat sheet, as there are on average 481 car accidents per month according to this data. Homicides and assaults are comparatively much lower at 5.8 and 70 per month. This provides perspective into areas that PG County police should focus on for the next year, and the PG County MVAs could also step in to provide more rigor in provisional license training. But on the crime side, police could look at this data and see that anywhere with a large magnitude of cars like parking garages would be likely places for car theft and could be more surveilant in those areas.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
