---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
crime_data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- crime_data |> mutate(violent_rate_2019 = violent2019/pop2019 * 100, property_rate_2019 = property2019/pop2019 * 100, violent_rate_2020 = violent2020/pop2020 * 100, property_rate_2020 = property2020/pop2020 * 100)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

ANSWER HERE
I know 100 is kinda lowball but it easily showed me outliers like Baltimore and Williamsport that were 1 whole unit higher than basically all other 130 cities. Crime is pretty low in most of Maryland per 100 people. Kinda interesting since media perception locally makes it feel like Maryland is dangerous or something sometimes. When really it has dangerous parts but most of MD has low murder rates. 

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate(pct_change = (violent_rate_2020- violent_rate_2019)/violent_rate_2019 * 100, (property_rate_2020- property_rate_2019)/property_rate_2019 *100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

ANSWER HERE
This was kind of illuminating, some cities had crazy crazy pct changes in different categories like Edmonston with 895% change in violent crime, Poolesville with a 600% in violent crime, and 320% property crime rate increase in Crisfield. Story in the data... I guess the outliers cause that's pretty significant changes jeez bro. I think the outliers are more interesting or pop out most.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE
MD violent crime has majorly decreased from 2019 to 2020

Although some cities in Maryland like Edmonston and Poolseville have had 800% and 600% increases in violent crime, for the most part it has gone down or is marginally decreased according to MD city data. Data was measured per 100 people which could potentially skew or dramaticize data, which is why these outliers don't take away from the fact that property crime and violent crime have mostly gone down from 2019 to 2020 in Maryland. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: Take the violent crime rates for 2020 and do three things:

Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile?
```{r}
md_cities_percentiles <- md_cities_rates |> mutate(percentil = ntile(violent_rate_2020, 100))
```
10 cities were above the 90th percentile. OC at the 100%, dang man...

Calculate the Interquartile Range of the 2020 violent crime rate using `summarize()` and R's `IQR()` function. The syntax is similar to how `mean()` works.
```{r}
interquartile_violent_2020 <- md_cities_rates |> summarize(IQR= IQR(violent_rate_2020))
```


Do the results change your best story idea from the previous assignment? Why or why not?
I guess my angle should've went the other way or maybe that crime went up a fair bit. I don't know what the heck interquartile range is but if that's the median for all of those cities then my angle on crime going down was def wrong.
