---
title: "Linear Regression"
output: html_document
---

---
title: "Identifying Schools with Biggest Differences in Math Test Scores, Part 2"
author: "Joseph Conaty"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```

### Introduction

We'll do what we did in class, but with a twist: instead of Montgomery County, Pa., you'll be working with data from Montgomery County, Md. Same basic idea, though: linear regression looking at free and reduced-price meals and math test scores.

Just as in the exercise, you will create a scatterplot to visualize the relationship between the FARMS (Free and Reduced-price Meal System) percentage and the percentage of students who met or exceeded math expectations. You'll then use linear regression to identify the schools that show the biggest differences between their actual and expected math test scores and evaluate the results of your model.

### Loading the Data

First, let's load the data into a data frame using the read_csv() function from the tidyverse library.

```{r}
schools <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/montgomery_md.csv")
```
### Task: Create a Scatterplot and Describe the Results (5 points)

First, let's create a scatterplot to visualize the relationship between the FARMS percentage and the percentage of students who met or exceeded math expectations. We'll use ggplot2 to create the plot and add a linear regression line.

```{r}
schools |>
  ggplot(aes(x = `FARMS Pct`, y = `% Met or Exceeded Math Expectations`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(x = "FARMS Percentage", y = "% Met or Exceeded Math Expectations", title = "Scatterplot of Math Expectations vs FARMS Percentage")
```
Based on what you see here, what's the relationship between FARMS percentage and math test scores? Compare this scatterplot to the one you made in the class activity.

ANSWER HERE
It's pretty hard to pull something guaranteed out of this data, it kind of looks like as farms percentage increases % met of math expectations decreases. However there are a significant and almost even amount of data points that say otherwise which is why it's hard to really pull something real out of this.

### Task: Linear Regression Model (7 points)

Next, you'll perform a linear regression analysis using the lm() function. We'll set the percentage of students who met or exceeded math expectations as the dependent variable, and the FARMS percentage as the independent variable. In other words, if we know the FARMS percentage for a school, how reliably can we predict its math score?

```{r}
model <- lm(`% Met or Exceeded Math Expectations` ~ `FARMS Pct`, data = schools)

summary(model)
```

Provide the model's output for the three statistics mentioned in the class activity.

ANSWER HERE
P value: < 0.00000000000000022
Adjusted r squared: 0.6734 
Coefficients: y = -0.59511x + 68.04878


Based on what you've learned so far, what does it tell us about the nature of the relationship between FARMS percentage and math scores? Is this model worth exploring further? Why or why not?

ANSWER HERE
Uh r squared is close to 100 so maybe there's something there, but also p value is so incredibly low I don't think whatever we do find is going to mean a strong relationship. I don't really know what to do with the slope intercept form equation formed by the coefficients or what it means.

### Task: Identifying Schools with Biggest Residuals (3 points)

Let's calculate and examine the residuals next. 

```{r}
schools$predicted <- predict(model)

# Calculate the residuals
schools <- schools |> 
  mutate(residual = `% Met or Exceeded Math Expectations` - predicted)

# Sort the data frame by the absolute value of the residuals in descending order
schools <- schools |>
  mutate(abs_residual = abs(residual)) |>
  arrange(desc(abs_residual))

```

Look at the absolute values of the residuals; what are the largest differences between the predicted result and the actual math scores? Pick two schools with the largest residuals. Find the zip codes where they are located and using the CensusReporter.org website, describe the most interesting demographic data from each zip code.

ANSWER HERE
Largest differences between predicted and actual are like 40 - 50% difference (I think it's in percentage), and the schools  with the two highest off top are Carderock and Cold Springs. Carderock took me to Bethesda, which is predominantly white and asian in terms of population, but interestingly the median age of residents is 61.8 and educational attainment is 90%. Academic weapons over there. Moving over to Cold Springs in Montgomery, which surprised me with how 74% of residents are white. WOW what the heck. Median household income is $250,000. I mean JEEEEZ. What is with this place man, 99.9% educational attainment. This place is different. Wow. 

## Task: Evaluate the model (5 points)

Describe what conclusions you'd draw from this model's results and how it would influence a story pitch about the school board member and math scores.

ANSWER HERE
Um I think the data I can kind of use and say sometimes FARMS can correlate with math scores or whatever, but what if I made a story all about how people can miscontrue data in a racist manner. And used the two counties we used in the answer here above to show how if you just saw these numbers raw it looks like white people are just insanely like smart or whatever but what it really is is an access to resources that a median household having $250,000 income can get. Tutors or programs and what have you that can lend to the child's success rate academically. And maybe throw in the school board member and stuff there and how the data is like comparing two things that almost have something in common but not quite and then jumping to conclusions that they relate.
